---
title: "Fc-OPG PK/PD in healthy post-menopausal women"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    css: src/styles/styles.css
    highlight: pygments
    theme: flatly
---

<hr><div class = spacersm></div>

```{r,echo=FALSE}
source("src/global.R")
```

<hr><div class = spacersm></div>

# Packages
```{r}
library(mrgsolve)
library(dplyr)
library(magrittr)
library(parallel)
library(ggplot2)
```

<hr><div class = spacer></div>

## Question
- What is the median change from baseline 14 days after 3 mg/kg SC dose?
- What is the probability that median CFB is greater than 40%?

<hr><div class = spacer></div>
# OPG model
```{r}
mod <- mread("opg", "model")

mod
```

Checkout the random effects structure
```{r}
revar(mod)
```


```{r}
mod %<>% mrgsolve:::collapse_omega() 
mod %<>% mrgsolve:::collapse_sigma()
```


We can render this in to an `Rmarkdown` document
```{r,eval=FALSE}
mrgsolve:::render(mod)
```

<hr><div class = spacer></div>

# Load the simulated posterior
```{r}
post <- readRDS("data/opgpost.RDS") %>% sample_n(1000)

post %>% head

param(mod)
```


For most of our examples, we are taking output directly
from a NONMEM run. This example is a little different:
We have changed the column names from `THETA1`, `THETA2` 
etc ... to `TVCL`, `TVVC` etc ...
```{r,eval=FALSE}
head(post)

inventory(mod,post)
```

```{r}
post <- mutate(post, TVVP1 = TVP1, TVVP2 = TVP2)
```

When working with results out of NONMEM, it is common to get
the `THETA` , `OMEGA` and `SIGMA` estimates all in a single 
row in a data frame.  

In order to get the information for `OMEGA` and `SIGMA` in to 
the problem, we need to go into this data and 
form matrices.

There are two specialized functions to help with this:

- `as_bmat` when the data is in block format
- `as_dmat` when the data is in diagonal format

```{r}
omegas <- as_bmat(post,"OMEGA")
sigmas <- as_bmat(post,"SIGMA")
omegas[[10]]
length(omegas)
is.list(omegas)
slice(post, 10)
omegas[[10]]
sigmas[[10]]
```


Here is a simple dosing data set to simulate 100 patients with 
3 mg/kg Fc-OPG SC x1
```{r}
sc3 <- expand.ev(ID=1:100, amt=210)
head(sc3)
```


We will get the observation design for the simulation 
through a `tgrid` object
```{r}
stime(tgrid(0, 12, 6) + 24)

des <- tgrid(end=-1,add=c(0,24*14))
des
```


When we do replicate simulation, it almost always pays off
to construct a function that will carry out one
replicate.


Arguments to the function are

- `i` the current simulation replicate
- `data` the dosing data set
- `des` the observation design


```{r}
sim <- function(i,data,des) {
  
  mod %<>% 
    param(slice(post,i)) %>%
    omat(omegas[[i]]) %>% 
    smat(sigmas[[i]])
  
  mod %>%
    Req(PDDV) %>%
    mrgsim(data=data,tgrid=des,obsonly=TRUE) %>% 
    mutate(irep=i) %>% 
    mutate(TVIC50=mod$TVIC50, TVQ1=mod$TVQ1, TVVMAX=mod$TVVMAX)
}
```


Also note 

- We update `param` with `ith` row from `post`
- We update `OMEGA` and `SIGMA` from the `ith` position 
in the appropriate list of matrices
- We will capture the current 


It is easy to test the function.  The 10th replicate will
look like this
```{r}
sim(10,sc3,des)
slice(post, 10)
```



Simulate reps
```{r}
options(mc.cores=8)
mcRNG()
set.seed(22223)
out <- lapply(1:1000, sim, data=sc3,des=des) %>% bind_rows

head(out)
```



Summarise

- First, get the baseline `NTX`
- Then calculate percent CFB


```{r}
sum <- 
  out %>%
  group_by(ID,irep) %>%
  mutate(BASE = first(PDDV), dDV = 100*(PDDV-BASE)/BASE) %>%
  ungroup
head(sum)
```



Filter down to week 2
```{r}
sum %<>% filter(time==336 & BASE >= 0)
```


Now, get the median
```{r}
summ <- 
  sum %>%
  filter(time==336) %>%
  group_by(irep) %>%
  summarise(med = median(dDV))

ggplot(summ, aes( x = med)) + 
  geom_histogram(alpha = 0.4, col = "white")

summ
```


<hr><div class = spacer></div>

## From the abstract

> Subsequent clinical trial simulations demonstrated that a single 3.0-mg/kg SC dose of Fc-OPG would be expected to produce, at 14 days post-dose, a median NTX percentage change from baseline of −45% (with a 95% prediction interval ranging from −34% to −60%)."

```{r}
ans <- signif(quantile(summ$med, c(0.5,0.025,0.975)),3)

paste0(ans[1], " (", ans[3],",",ans[2],")")
```

Median week-2 change from baseline with 95% interval


Probability that median cfb > 40%
```{r}
mean(summ$med < -40)
```



Plot the distribution of the week-2 change from baseline
```{r}
ggplot(summ, aes(x=med)) + 
  geom_histogram(fill="darkslateblue", col="grey")
```


<hr><div class = spacer></div>

# Sensitivity analysis

We are already set up to do the sensitivity analysis

```{r}
par <- dplyr::distinct(out,irep,TVIC50,TVQ1,TVVMAX)
head(par)
sens <- left_join(summ,par)
head(sens)
```

`TVIC50`
```{r}
ggplot(sens, aes(TVIC50, med)) + 
  geom_point(col="darkslateblue") + 
  geom_smooth(method="loess",col="red",lwd=2)
```

`TVVMAX`
```{r}
ggplot(sens, aes(TVVMAX, med)) + 
  geom_point(col="darkslateblue") +
  geom_smooth(method="loess",col="red",lwd=2)
```


`TVQ`
```{r}
ggplot(sens, aes(TVQ1, med)) + 
  geom_point(col="darkslateblue") +
  geom_smooth(method="loess",col="red",lwd=2)

```

