---
title: "VPC - meropenem PK model"
author: "Metrum Research Group, LLC"
date: ""
output: 
  html_document:
    css: src/styles/styles.css
    highlight: pygments
    theme: flatly
---


```{r, message = FALSE}
source("src/global.R")
source("src/functions.R")
library(dplyr)
library(mrgsolve)
library(readr)
library(magrittr)
library(ggplot2)
library(parallel)
```


# Introduction

- "Population Pharmacokinetic Analysis and Dosing Regimen Optimization of 
  Meropenem in Adult Patients"
    - Li et al. J Clin Pharmacol 2006
- Meropenem is broad-spectrum carbapenem antibiotic
    - Efficacy related to time above MIC
- IV dosing every 8 hours by infusion or bolus
    - bolus over 3 to 5 minutes
    - infusion over 15 to 30 minutes
- Authors are interested in seeing if a longer infusion duration 
  can increase time above MIC

# Helper functions
```{r}
qt <- function(x,y) unname(quantile(x,prob=y/100))
lo <- function(x) qt(x,5)
hi <- function(x) qt(x,95)
med <- function(x) qt(x,50)
loci <- function(x) qt(x,2.5)
hici <- function(x) qt(x,97.5)
col1 <- "steelblue"
col2 <- "firebrick"
```

# Read and fix up the data set
```{r}
data <- 
  read_csv("data/Simulated_DatasetMeropenem.csv", na = '.')  %>% 
  mutate(CMT=1, DUR = AMT/RATE)
```



Derive a column that describes the infusion duration for each subject
```{r}
data %<>% 
  group_by(ID) %>% 
  mutate(DUR = first(DUR[!is.na(AMT)])) %>% 
  ungroup

data <- data %>% mutate(DUR = round(DUR,1))
```

Look at distinct values of `CMT`, `EVID`, `DUR` in `data`

```{r}
count(data,CMT,EVID,DUR)
```

Derive two data frames

- One with observations only
- One with doses only

```{r}
obs <- filter(data, EVID==0)
head(obs)
dose <- filter(data, EVID==1)
```


Plot observed data
```{r}
ggplot(data=obs, aes(TIME,DV)) + 
  geom_point() +
  scale_y_continuous(trans = "log", breaks = logbr())
```

Plot observed data by `DUR`
```{r}
ggplot(data=obs, aes(TIME,DV)) + 
  geom_point() +
  scale_y_continuous(trans = "log", breaks = logbr()) +
  facet_wrap(~DUR) + xlim(0,8)
```

# Load the meropenem model

```{r}
mod <- mread("meropenem", "model")
see(mod)
```

This model looks a little different because we got it 
off of DDMoRe model repository.


# Set up a simulation time grid for the VPC

We want 
- Hourly observations from time of first dose to 8 hours
- Observations every 0.1 hours between 0 and 3 hours 

```{r}
des1 <- tgrid(0,3.1,0.1)
des2 <- tgrid(0,8,1)
des <- c(des1,des2)
des
```


A function to do the (replicate) simulation

Arguments

- `i` the replicate number

Returns simulated data set 

- time as in `des`
- `DUR` the infusion duration 
- `TIME > 0` and `Y > 0`
- Labeled with replicate number

```{r}
simvpc <- function(i) {
  mod %>% 
    data_set(dose) %>%
    carry_out(DUR) %>%
    obsonly %>%
    mrgsim(tgrid=des) %>% 
    filter(TIME > 0 & Y > 0) %>%
    mutate(irep = i)
}
```

Simulate

- 100 iterations
- Use mclapply
- Bind into a single data frame

```{r}
niter <- 100
out <- mclapply(1:niter, mc.cores=8, simvpc) %>% bind_rows
```


# Summarize simulated data
```{r}
sum1 <- 
  out %>% 
  filter(Y > 0) %>%
  group_by(DUR,irep,TIME) %>%
  summarise(med=med(Y), lo=lo(Y), hi=hi(Y), N=n())
```


```{r}
sum2 <- 
  sum1 %>%
  group_by(DUR,TIME) %>% 
  summarise(medlo = loci(med), medmed = med(med), medhi = hici(med),
            lolo =  loci(lo),  lomed  = med(lo),  lohi  = hici(lo),
            hilo =  loci(hi),  himed  = med(hi),  hihi  = hici(hi))
```

```{r}
p1 <- 
  ggplot(data=sum2) + 
  geom_ribbon(aes(TIME,ymin=medlo, ymax = medhi),alpha=0.3,fill=col1) + 
  geom_ribbon(aes(TIME,ymin=lolo, ymax=lohi),alpha=0.3,fill=col1) + 
  geom_ribbon(aes(TIME,ymin=hilo, ymax=hihi),alpha=0.3,fill=col1) + 
  geom_point(data=obs, aes(TIME,DV),col=col2) +
  geom_line(aes(TIME,y=medmed), lwd=1,col=col1) +
  geom_line(aes(TIME,y=lomed), lwd=1,col=col1) +
  geom_line(aes(TIME,y=himed), lwd=1,col=col1) +
  scale_y_continuous(trans="log", breaks=10^seq(-5,5)) +
  facet_wrap(~DUR)

```


```{r, fig.width=8}
p1
```

  
# Summarize observed data and add to plot
```{r}
obs1 <- 
  obs %>% 
  filter(DV > 0) %>%
  group_by(DUR,TIME) %>%
  summarise(med=med(DV), lo=lo(DV), hi=hi(DV), N=n())
```

```{r, fig.width=8}
p1 + 
  geom_line(data=obs1,aes(TIME,y=med),lty=2, lwd=1) +
  geom_line(data=obs1,aes(TIME,y=lo), lty=2, lwd=1) +
  geom_line(data=obs1,aes(TIME,y=hi), lty=2, lwd=1)
```



