---
title: "Fc-OPG PK/PD in healthy post-menopausal women"
author: ""
date: ""
output: 
  html_document:
    css: src/styles/styles.css
    highlight: pygments
    theme: flatly
---

<hr><div class = spacersm></div>

```{r,echo=FALSE}
source("src/global.R")
```

# Packages
```{r}
proj <- "models"
library(mrgsolve)
library(dplyr)
library(magrittr)
library(parallel)
library(ggplot2)
```

<hr><div class = spacer></div>

## Question
- What is the mean change from baseline 14 days after 3 mg/kg SC dose?
- What is the probability that mean CFB is greater than 40%?

# Code the OPG model


<hr><div class = spacer></div>
# OPG model

```{r load and checkout opg model}

```

Checkout the random effects structure
```{r}

```

Collapse `$OMEGA` and `$SIGMA`
```{r collapse_omega_sigma}

```


We can render this in to an `Rmarkdown` document
```{r,eval=FALSE}

```

# Load the simulated posterior

- `data/opgpost.RDS`
- Sample 1000 sets

```{r read opgpost file }
#set.seed()
post <- NULL
```


For most of our examples, we are taking output directly
from a NONMEM run. This example is a little different:
We have changed the column names from `THETA1`, `THETA2` 
etc ... to `TVCL`, `TVVC` etc ...
```{r,check out posterior}

```


When working with results out of NONMEM, it is common to get
the `THETA` , `OMEGA` and `SIGMA` estimates all in a single 
row in a data frame.  

In order to get the information for `OMEGA` and `SIGMA` in to 
the problem, we need to go into this data and 
form matrices.

There are two specialized functions to help with this:

- `as_bmat` when the data is in block format
- `as_dmat` when the data is in diagonal format

```{r}
omegas <- NULL
sigmas <- NULL
```


Here is a simple dosing data set to simulate 100 patients with 
3 mg/kg Fc-OPG SC x1
```{r}
sc3 <- NULL
head(sc3)
```


We will get the observation design for the simulation 
through a `tgrid` object

Let's only get observations at baseline and at week-2

```{r tgrid object}
des <- numeric(0)
stime(des)
```


When we do replicate simulation, it almost always pays off
to construct a function that will carry out one
replicate.


Arguments to the function are

- `i` the current simulation replicate
- `data` the dosing data set
- `des` the observation design

Workflow
- Update parameters, omega and sigma
- Simulate from `data` using `des`
  - Get observations only
  - label the replicate number with `irep`
  - Also capture `TVIC50`, `TVQ1`, and `TVVMAX`
  
Return
 - A data frame
 
```{r}
sim <- function() {
  
}
```


Also note 

- We update `param` with `ith` row from `post`
- We update `OMEGA` and `SIGMA` from the `ith` position 
in the appropriate list of matrices
- We will capture the current 


It is easy to test the function.  The 10th replicate will
look like this
```{r test the sim function}

```



Simulate reps

- Simulate with `mclapply` with 8 cores
- 1000 replicates
- Return a single data frame
- Rember to set a seed

```{r, eval = FALSE}
options(mc.cores=8)
mcRNG()
#set.seed()
out <- NULL
```



Summarise

- For each individual
  - First, get the baseline `NTX`
  - Then calculate `percent` CFB



```{r}
sum <- 
  out %>%
  group_by() %>%
  mutate(BASE = NULL, dDV = NULL) %>%
  ungroup
```



Filter down to week 2
```{r}

```


Now, get the median change from baseline for each replicate
```{r}
summ <- NULL
```


<hr><div class = spacer></div>

## From the abstract

> Subsequent clinical trial simulations demonstrated that a single 3.0-mg/kg SC dose of Fc-OPG would be expected to produce, at 14 days post-dose, a median NTX percentage change from baseline of −45% (with a 95% prediction interval ranging from −34% to −60%)."

```{r}
ans <- NULL
```

Median week-2 change from baseline with 95% interval


Probability that median cfb > 40%
```{r}

```



Plot the distribution of the week-2 change from baseline
```{r}

```


<hr><div class = spacer></div>

# Sensitivity analysis

We are already set up to do the sensitivity analysis

- `TVIC50`, `TVQ1`, and `TVVMAX`

```{r, eval = FALSE}
par <- dplyr::distinct(out,irep,...)
head(par)
```

Now join the parameters
```{r, eval = FALSE}
sens <- left_join(summ,par)
head(sens)
```

`TVIC50`
```{r, eval = FALSE}
ggplot(sens, aes(TVIC50, med)) + 
  geom_point(col="darkslateblue") + 
  geom_smooth(method="loess",col="red",lwd=2)
```

`TVVMAX`
```{r, eval = FALSE}
ggplot(sens, aes(TVVMAX, med)) + 
  geom_point(col="darkslateblue") +
  geom_smooth(method="loess",col="red",lwd=2)
```


`TVQ`
```{r, eval = FALSE}
ggplot(sens, aes(TVQ1, med)) + 
  geom_point(col="darkslateblue") +
  geom_smooth(method="loess",col="red",lwd=2)

```

