
---
title: "Introduction to <grn>mrgsolve</grn>"
author: PAGE 2018 Workshop
date: Metrum Research Group, LLC
knit_root_dir: "../../"
output: 
  slidy_presentation:
    number_sections: true
    css: [../../src/styles/slidystyles.css,../../src/styles/colors.css]
  html_document:
    theme: united
    number_sections: true
    toc: true
    css: [src/styles/htmlstyles.css,src/styles/colors.css]
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, cache.path="cache/deck_00/")
```


```{r, include = FALSE}
source("src/global.R")
library(mrgsolve)
library(dplyr)
library(knitr)
library(lattice)

tryit_file <- "workbook.Rmd"
```

# People

- Michelle Johnson
- Matthew Riggs
- Bill Knebel

> - Kyle Baron
> - The  participants ... thank you for coming!

# Cloud-based computer access via Metworx

- Each participant may access an 8 core cloud instance 
using Metworx
- All software (R packages) needed for the workshop is pre-installed
- We will use Rstudio and the R package mrgsolve for the course exercises
- System requirements: Laptop with wifi and a browser 

# Accessing Metworx

<h3><grn>https://metworx-eu-central-stage.metworx.com</grn></h3>

<h3>
- Sign in to Metworx
    - <blu>username:</blu> page2018-<red>n</red>@metrumrg.com
    - <blu>password</blu> PageConf2018!

<hr>

- Sign in to Rstudio
    - Click on Rstudio
    - <blu>username:</blu> page2018
    - <blu>password:</blu> PageConf2018!
</h3>

# Check out the repo

```{r, eval = FALSE}
username <- "mrgsolve-sp18"
password <- "a45f7aa3d8f8fa53b25d2db780f6077ccba80925"

#############################

.libPaths("/data/page-Rlibs")

if(!require("git2r")) install.packages("git2r")

library(git2r)

.repo_addr <- "https://github.com/metrumresearchgroup/page-2018-mrgsolve"
.where <- "page-2018-mrgsolve-workshop"
.branch <- "master"

if(!file.exists(.where)) dir.create(.where)

repo <- clone(
  .repo_addr, 
  branch=.branch,
  credentials=cred_user_pass(username,password),
  local_path=.where
)
```


```{r setup, echo = FALSE, message = FALSE}
mod <- mread_cache("pk1", modlib()) %>% 
  update(end = 192, delta = 0.2) %>% Req(CP)
data(exidata)
data <- filter(exidata, ID <= 10)
set.seed(1222)
```

# Why “Open Science”?

- What is this concept of “open” ?
- How & why is it beneficial to the growth of our community 
- Examples:
    - open-source software: SBML, mrgsolve, Shiny etc., STAN/Torsten(?)
    - open model repositories: Biomodels, github, DDmore. . .
    - Publications, presentations, regulatory docs:
        - good (model is described with parameter estimates),
        - better (model mathematics provided with parameter estimates),
        - best (executable code)
    - PKsim / MOBI
    - Osmosis Foundation

# How we implement our science

- Overview of Metworx system
    - why share when you don’t have to (Performance),
    - how to easily share when you want to (Envision)

# Ideal computing

<img src = "../../src/images/cloud.pdf" width = 800px></img>

# What we might want

<img src = "../../src/images/shinyPic.pdf" width = 800px></img>

# Our workflow (Metworx)

<img src = "../../src/images/metworxFlow.pdf" width = 800px></img>



---

```{r, echo = FALSE}
mod %>% ev(amt = 100) %>% mrgsim(end = 72) %>% plot()
```

---

```{r, echo = FALSE}
mod %>% ev(amt = 100, ii = 24, addl = 4) %>% mrgsim() %>% plot()
```

---

```{r, echo = FALSE}
mod <- mread_cache("popex", modlib()) %>% Req(DV)
data <- expand.ev(amt = 300, ii = 24, addl = 4, ID = seq(20))
mod %>% data_set(data) %>% mrgsim() %>% plot()
```

---

```{r, echo = FALSE}
hmod <- mrgsolve:::house() %>% Req(DV,RESP) %>% update(end = 244)
hdat <- mutate(data, time = 24)
mrgsim(hmod, data = hdat, omega = dmat(0.1,0.1,0.1,0.1)) %>% plot()
```

---

```{r, echo = FALSE}
idata <- data_frame(VC = c(20,60,180))
hmod %>% mrgsim(idata = idata,  end = 24*28, delta = 0.1,
                events = ev(amt = 100, ii = 24, addl = 28)) %>%
  plot(DV~.)
```


---

```{r, echo = FALSE}
mod <- mread_cache("conway", "model")
e <- ev(amt = 1, evid = 8, time = 180) + ev(amt = 0, evid = 8, time = 2.5*365)
idata <- data_frame(L_0 = c(1,3,5,7,10))
out <- mrgsim(mod, events = e, idata = idata, end = 6*365, delta = 1)
out@data <- mutate(out@data, time = time - 180)
plot(out, L+logV~(time/365), xlab = "Time (year)")
```


# About `mrgsolve` 

- `R` package for simulation from ODE-based models
    - Free, OpenSource, GitHub, CRAN
- Language
    - Models written in `C++` inside model specification format
    - General purpose solver: `ODEPACK` / `DLSODA` (`FORTRAN`)
    - Simulation workflow in `R`
- Hierarchical (population) simulation
    - `ID`, $\eta$, $\varepsilon$
- Integrated PK functionaility
    - Bolus, infusion, `F`, `ALAG`, `SS` etc, handled under the hood
    - 1- and 2-cmt PK models in closed-form
- Extensible using `R`, `C++`, `Rcpp`, `boost`, `RcppArmadillo`
- `R` is it's natural habitat

# `mrgsolve` started as `QSP` modeling tool
- Motivation: large bone/mineral homeostatsis model (CaBone)
- History using
    - Berkeley Madonna
    - WinBUGS
    - NONMEM (attempted)
- 2010: write `R` front end to `deSolve`
- 2012: write `C++` interface to `DLSODA`
- Develop dosing / event capability
- More recently, expose functionality provided by
    - `Rcpp` - vectors, matrices, functions, environments, random numbers 
    - `boost` - numerical tools in `C++`
    - users' own `C++` code (functions, data structures, classes)
- Translator from `SBML` to `mrgsolve` using `R` bindings to `libSBML`   


# Orientation

- https://CRAN.R-project.org/package=mrgsolve

- GitHub site: https://github.com/metrumresearchgroup/mrgsolve

- mrgsolve website: https://mrgsolve.github.io

- User Guide: https://mrgsolve.github.io/user_guide

# What we will cover today

1. Three basic workflows
1. Loading the model into R
1. Event objects
1. Data sets
1. Model specification - code together
1. This and that
1. Applied examples
1. Whatever you ask about

<h3><grn>Emphasis is on getting you running your own simulations today.</grn></h3>
