
---
title: "Introduction to <grn>mrgsolve</grn><BR>Models and Events"
author: PAGE 2018 Workshop
date: Metrum Research Group, LLC
knit_root_dir: "../../"
output: 
  slidy_presentation:
    number_sections: true
    css: [../../src/styles/slidystyles.css,../../src/styles/colors.css]
  html_document:
    theme: united
    number_sections: true
    toc: true
    css: [src/styles/htmlstyles.css,src/styles/colors.css]
---


```{r, include = FALSE}
source("src/global.R")
library(mrgsolve)
library(dplyr)
library(knitr)
library(lattice)
tryit_file <- "workbook.Rmd"
```

```{r setup-1, echo = FALSE, message = FALSE}
mod <- mread_cache("pk1", modlib()) %>% 
  update(end = 192, delta = 0.2) %>% Req(CP)
data(exidata)
data <- filter(exidata, ID <=10)
set.seed(1222)
```

# A basic simulation with mrgsolve
```{r}
mod %>% ev(amt = 100, ii = 24, addl = 3) %>% mrgsim() %>% plot()
```


# A basic simulation with mrgsolve

```{r, eval = FALSE}
mod %>% ev(amt = 100, ii = 24, addl = 3) %>% mrgsim() %>% plot()
```

- <red>mod</red>: the model object
    - Ok ... where did `that` come from?
- <blu>ev(amt = 100, ...)</blu> : the intervention
    - An `event` in this example
- <grn>mrgsim()</grn>: actually do the simulation
- <orng>plot()</orng>: do something with the simulation
    - `plot`, `mutate`, `as_data_frame` etc ... 


> - <red>model</red> `%>%` <blu>intervention</blu> `%>%` <grn>Go!</grn> `%>%` <orng>take-a-look</orng> 

# What's coming ...

> - `model %>% intervention %>%` <alrt>options</alrt> `%>% Go! %>% ...`

> - `model %>% intervention %>%` <orng>population</orng> `%>% Go! %>% ...`

> - `model %>%` <purp>data-set</purp> `%>% Go! %>% ...`

> - where <purp>data-set</purp> = `intervention + population`

> - For now, let's get this part down 

> - <red>model</red> `%>%` <blu>intervention</blu> `%>%` <grn>Go!</grn> `%>%` <orng>take-a-look<orng> 


# Why do we use `%>%` ?

What happens first in this operation?

```{r, eval = FALSE}
mean(sqrt(seq(4)))
```

# Pipelines

```{r, eval = FALSE}
mean(sqrt(seq(4)))
```

```{r, eval = FALSE}
4  %>% seq() %>% sqrt() %>% mean()
```

Better.

```{r, eval = FALSE}
4  %>% 
  seq(.) %>% 
  sqrt(.) %>% 
  mean(., na.rm = TRUE)
```

```{r,eval = FALSE}
mod %>% some_intervention() %>% simulate() %>% post_process()
```


# The model object

<red>mod</red><blk>el</blk> `%>% ... `

> - I (almost) always call the model object <red>mod</red> in the documention / examples

> - All the information about the model we need to know to run the simulation

> - <grn>Distinct</grn> from the intervention, the population, the summary

> - But the model <orng>does</orng> know  about output time, random effects





# Take a look: overview

```{r, comment = ""}
mod
```

# Take a look: parameters (<red>really important</red>)
```{r, comment = ""}
param(mod)
```

> - Parameters get a name
> - Names and number of parameters gets fixed at compile time
> - All parameters have a value
> - Value can be modified after compile time

# Take a look: compartments
```{r, comment = ""}
init(mod)
```

> - Every compartment gets a name
> - Every compartment gets an initial condition

# Where did <red>mod</red> come from?

```{r,eval = FALSE}
mod <- mread("simple", "model")
```


```{r, eval = FALSE}
mod <- mread("<model-name>", "<project-directory>")
```

- By default mrgsolve looks for the code in the file
    - <grn>model-name.cpp</grn> in
    - <blu>project-directory</blu>

- <red>mread demo</red>

```{r, echo = FALSE}
Sys.sleep(2)
```


# Read in a model object with caching

First time to read
```{r}
mod <- mread_cache("simple", "model")
```

Next time to read

```{r}
mod <- mread_cache("simple", "model")
```

# Oops 
```{r}
mod <- try(mread("simple"))

mod
```


# Internal model library

Quiz:

```{r, eval = FALSE}
mod <- mread("<first-argument>", "<second-argument>")
```

# Internal model library

Quiz:

```{r, eval = FALSE}
mod <- mread("<first-argument>", "<second-argument>")
```

```{r, eval = FALSE}
mod <- mread("<first-argument>", modlib())
```

```{r}
modlib()
```


# Internal model library
```{r}
mod <- mread("effect", modlib())
```

```{r}
mod
```


# Inline model specification

We <red>haven't covered</red> the specifics of coding a model yet

```{r}
code <- '
$PARAM CL = 1, V = 20
$PKMODEL cmt = "CENT"
'
```

```{r, eval = FALSE}
mod <- mcode("dont_do_this", code)
```

```{r, eval = FALSE}
mod <- mcode_cache("seriously_dont", code)
```

> - <font class = "red">Question:</font>  Why is this a bad idea?

# <red>Your turn</red>

- File name: <grn>exercises/warm_up.R</grn>

- Section name: <blu>Warm Up</blu>


# Event objects

- `model %>%` <red>intervention</red> `%>% Go! %>% take-a-look`

- Event object  = quick / easy way to implement dose or other intervention


```{r}
e <- ev(amt = 100) 

e
```

> - Defaults: `time`, `evid`, `cmt`

# Three ways to invoke

<blk>Inline</blk>

```{r, eval = FALSE}
mod %>% ev(amt = 100) %>% mrgsim()
```

<blk>Object via pipeline</blk>

```{r, eval = FALSE}
e <- ev(amt = 100)

mod %>% ev(e) %>% mrgsim()
```

<blk>As argument</blk>

```{r, eval = FALSE}
mod %>% mrgsim(events = e)
```


# What to include in `ev(...)`

- <grn>time</grn> event time
- <grn>cmt</grn> Event compartment
- <grn>amt</grn> Dose amount
- <grn>ii</grn> Inter-dose interval
- <grn>addl</grn> Additional doses to administer
- <grn>rate</grn> Infusion rate
- <grn>ss</grn> Set to 1 to advance to steady state
- <grn>evid</grn> Event id
- <grn>ID</grn> Subject ID (use multiple ids - ID=1:10)

# Interventions and corresponding `evid`

-  <grn>Bolus</grn> dosing (`evid` 1, with `rate`==0)
-  Zero order <grn>infusion</grn> (`evid` 1, with `rate` > 0)
-  <grn>Other</grn> type event (`evid` 2)
    -  This also forces solver reset
-  Compartment <grn>reset</grn> (`evid` 3)
-  <grn>Reset</grn> and dose (`evid` 4)
-  <grn>Replace</grn> the amount in a specific compartment (`evid` 8)

# Create complex events - 1

What's going to happen?

```{r, eval = FALSE}
e1 <- ev(amt = 200) 

e2 <- ev(amt = 100, time = 24, ii = 24, addl = 4)

c(e1, e2)
```

# Create complex events - 1

What's going to happen?

```{r, echo = FALSE}
e1 <- ev(amt = 200) 
e2 <- ev(amt = 100, time = 24, ii = 24, addl = 4)
e <- c(e1,e2)
mod %>% 
  ev(e) %>% 
  mrgsim(end = 96) %>% plot(CP~.)
```



# Create complex events - 1

Combine two events

```{r}
e1 <- ev(amt = 200) 

e2 <- ev(amt = 100, time = 24, ii = 24, addl = 4)

c(e1, e2)
```

# Create complex events - 2

What's going to happen?

```{r, eval = FALSE}
e1 <- ev(amt = 200, ii = 12, addl = 2) 

e2 <- ev(amt = 100, ii = 24, addl = 4)

seq(e1, e2)
```

# Create complex events - 2

What's going to happen?

```{r, echo = FALSE}
e1 <- ev(amt = 200, ii = 12, addl = 2) 
e2 <- ev(amt = 100, ii = 24, addl = 4)
e <- seq(e1, e2)
mrgsim(mod, events = e, end = 180) %>% plot(CP ~ .)
```


# Create complex events - 2

Put two events in a sequence

```{r}
e1 <- ev(amt = 200, ii = 12, addl = 2) 

e2 <- ev(amt = 100, ii = 24, addl = 4)

seq(e1, e2)
```


# Create complex events - 3

What is going to happen?

```{r, eval = FALSE}
e1 <- ev(amt = 200) 

e2 <- ev(amt = 100, ii = 24, addl = 4)

seq(e1, wait = 36, e2)
```

# Create complex events - 3

What is going to happen?

```{r, echo = FALSE}
e1 <- ev(amt = 200) 
e2 <- ev(amt = 100, ii = 24, addl = 4)
e <- seq(e1, wait = 36, e2)
mrgsim(mod, events = e, end = 180) %>% plot(CP ~ .)
```

# Create complex events - 3

Wait before starting the next part of the regimen

```{r}
e1 <- ev(amt = 200) 

e2 <- ev(amt = 100, ii = 24, addl = 4)

seq(e1, wait = 36, e2)
```

# Simulate

```{r, echo = FALSE, message = FALSE}
mod <- mread_cache("effect", modlib())
```


<hr>

`model %>% intervention %>%` <grn>Go!</grn>

<hr>



```{r}
e <- ev(amt = 100)

mod %>% ev(e) %>% mrgsim()
```


```{r, eval = FALSE}
mrgsim(mod, events = e)
```

# What would you like to "fix" in this plot?
```{r,echo= FALSE}
mod <- update(mod, end = 18, delta = 6)
```

```{r}
mod %>% ev(amt = 100) %>% mrgsim() %>% plot(CP~time)
```

# Simulation end time
```{r}
mod %>% ev(amt = 100) %>% mrgsim(end = 48) %>% plot(CP~time)
```

# Simulation time step
```{r}
mod %>% ev(amt = 100) %>% mrgsim(end = 48, delta = 0.1) %>% plot(CP~time)
```


# Simulation time grid

-  To form a sequence of times
    - <grn>start</grn> usually `0`
    - <grn>end</grn> time of last observation
    - <grn>delta</grn> <blk>output</blk> time step
-  Additional times to simulate
    - <purp>add</purp> ad-hoc numeric vector
    
```{r}
stime(mod)
```


# <red>Your turn</red>

- File name: <grn>exercises/z-pak.R</grn>

- Section name: <blu>Z-Pak</blu>


# Event objects are just data frames

```{r}
as.data.frame(e1)
```

- We will use a <blk>data_set</blk> later on for populations

- Event objects are convenient
  - Constructor
  - Operations

# We're stil working on this setup

<hr>

<red>model</red> `%>%` <blu>intervention</blu> `%>%` <grn>Go!</grn> `%>%` <orng>take-a-look</orng> 

<hr>

<red>model</red>:

- Load a model with `mread` or `mread_cache`
- Use the internal library with `mread("<model-name>", modlib())`
- Check model parameters with `param(mod)`
- Check model initial conditions with `init(mod)`
- View model code with `see(mod)`

<blu>intervention</blu>:

- `ev(...`): `amt`, `cmt`, `time`, `ii`, `addl`, `rate`
- Different ways to combine event objects



# Working with simulated output

<hr>

`model %>% intervention %>% Go! %>%` <orng>take-a-look</orng> 

<hr>

# Plot

```{r}
mod <- mread_cache("effect", modlib())

mod %>% ev(amt = 100) %>% mrgsim() %>% plot()
```

# Plot

```{r}
mod %>% ev(amt = 100) %>% mrgsim() %>% 
  plot(CP + EFFECT~., col = "firebrick")
```

# Pipeline to dplyr functions

```{r}
mod %>% mrgsim() %>% mutate(arm = 1)
```


# Update

On the fly

```{r, eval = FALSE}
mod %>% update(end = 72) %>% mrgsim()
```

Persistent 

```{r, eval = FALSE}
mod <- update(mod, end = 72)
```

mrgsim will call `update` for you (on the fly)
```{r, eval = FALSE}
mod %>% mrgsim(end = 72)
```


# What else can I update?
* Time
    - `start`, `end`, `delta`, `add`
* Parameters and compartment initial values
* Solver settings
    * `atol`, `rtol`
    * `hmax`, `maxsteps`, `mxhnil`, `ixpr`
* `$OMEGA`, `$SIGMA`
* `tscale` (rescale the output time)
* `digits`

```{r, eval = FALSE}
mod %>% update(rtol = 1E-12) %>% ...
```

```{r, eval = FALSE}
mod %>% mrgsim(rtol = 1E-12) %>% ...
```

# Updating parameters

- Demo

```{r,eval = FALSE}
mod <- param(mod, CL = 1.5)
```


# Controlling output - <grn>request</grn>

```{r}
mod %>% mrgsim() %>% head(n = 2)
```

```{r}
mod %>% Req(CP) %>% mrgsim() %>% head(n = 2)
```


# Controlling output - <grn>obsonly</grn>
```{r}
mod %>% ev(amt=1) %>% mrgsim() %>% head(n = 2)
```

```{r}
mod %>% ev(amt=1) %>% obsonly() %>% mrgsim() %>% head(n = 2)
```




